\chapter{Conclusions}
In conclusion let's summarize the results we obtained. Considering the compact models, we have that the flow based formulations, in particular flow 1 and flow 2, perform better than the MTZ and the time based formulations. Looking to the Performance profile we can conclude that flow 1 is the best compact model.
Moving to the exact algorithms, here we use the base TSP model, with different strategies to add the subtour eliminator constraints during (with callbacks) or after (loop method) the execution of CPLEX. In general we have that this type of algorithms works better than a simple compact model. Again looking to the performance profile we can establish that the combination of the UserCut and the Heuristic Callback based on the Generic Callback is the best strategy.
In the matheuristic chapter we analyze some heuristic solutions that are based on CPLEX and we reach the conclusion that hard fixing is the best matheuristic.
Finally in the stand-alone heuristics chapter we explore several strategies that don't provide an exact solution and are independent from CPLEX. We conclude that VNS is probably the best performing heuristic and simulated annealing and tabu search can be good solutions too. \\

\noindent So if the goal is to resolve a relatively small instance of the TSP, we suggest to use CPLEX plus heuristic callback plus user cut. Otherwise if we have to resolve a big instance or we are not interested to obtain the exact solution and we have a time limit that is very important to respect we suggest VNS. 


\bibliographystyle{ieeetr}
\bibliography{biblio}
